State of the Practice for Lattice Boltzmann Method Software

### 2022-03-06 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
reviewed abstract and section 3 - ranking projects..

All of the 24 software packages that were tested have installation instructions.
Most have their installation instructions located in one place, often in an
instruction manual or on a web-page. Sometimes, like with Ludwig, incomplete
installation instructions are found on a home page, with more detailed
instructions located on another web-page, or within the documentation.
Installability of the software, and maintainability and correctness of the
instructions themselves, are improved if all the instructions are in one
location.

@@

Tutorials are available for 19 of the software packages. Generally they are
linearly written and easy to follow. However, only eight tutorials provide an
expected output. It is not possible to verify the correctness of the output of
the software packages that are missing this key information. In these cases the
user may need to assume correctness if there are no visible errors.  A
particularly  detailed tutorial that include expected output is provided by
\href{https://www.walberla.net/doxygen/index.html} {waLBerla}.

@@

Of the 14 software packages that installed correctly and also have tutorials,
four (pyLBM, ESPResSo++, LIMBES, Ludwig) broke during tutorial testing. All of
these instances resulted in an error message being displayed. One error (pyLBM)
was due to a missing tutorial dependency, another (Ludwig) was due to an invalid
command despite following the tutorial, and the final two errors were vague
execution errors. Of the four broken tutorial instances, only the missing
dependency one was recoverable.

@@

Issue tracking is used in 23 software packages, 15 of which use GitHub or GitLab
to host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub/GitLab have the majority of
their issues closed, and only three (laboetie, lettuce, Sailfish) have less than
50 percent of their issues closed. All of the top five overall ranked packages
have most of their issues closed. Alive packages (11 use issue tracking) have
64\% of their issues closed, while dead packages (3 use issue tracking) have
71\% of their issues closed. This information is presented in
Table~\ref{gitrepodata}. With respect to the host for the git repository, 13
packages use GitHub and two (Palabos, waLBerla) use GitLab. Of the other
packages, one package (SunlightLB) uses CVS for issue tracking and version
control, and seven packages do not appear to use any issue tracking system.

@@

Software package artifacts were reviewed for the identification of a specific
development model, like a waterfall of agile development model, and the presence
of documentation recording the development process and standards. They were also
reviewed for the identification of the development environment, and the presence
of release notes. The packages tended to not explicitly use well-known
development models. This was also noted in the interviews with developers
(Section~\ref{Sec_CompareMethodologies}). The development teams of these
packages are fairly small and easily organized without the need for such
processes. Seven of the software packages did have some artifacts outlining the
general development process, how to contribute, and the status of the package or
its components. Eight of the packages have artifacts that note the development
environment. While this information could help developers, and would improve
transparency, the small close-knit nature of the development teams make
explicitly publicly specifying this information practically unnecessary. Version
release notes were found in nine of the software packages.

>>>>

All of the 24 software packages that were tested have installation instructions.
Most have their installation instructions located in one place, often in an
instruction manual or on a web-page. Sometimes, like with Ludwig, incomplete
installation instructions are found on a homepage, with more detailed
instructions located on another web-page, or within the documentation.
Installability of the software, and maintainability and correctness of the
instructions themselves, are improved if all the instructions are in one
location.

@@

Tutorials are available for 19 of the software packages. Generally they are
linearly written and easy to follow. However, only eight tutorials provide an
expected output. It is not possible to verify the correctness of the output of
the software packages that are missing this key information. In these cases the
user may need to assume correctness if there are no visible errors.  A
particularly  detailed tutorial that includes expected output is provided by
\href{https://www.walberla.net/doxygen/index.html} {waLBerla}.

@@

Of the 14 software packages that installed correctly and also have tutorials,
four (pyLBM, ESPResSo++, LIMBES, Ludwig) broke during tutorial testing. All of
these instances resulted in an error message being displayed. One error (pyLBM)
was due to a missing tutorial dependency, another (Ludwig) was due to an invalid
command despite following the tutorial, and the final two errors were vague
execution errors. Of the four broken tutorial instances, only the one due to a missing dependency was recoverable.

@@

Issue tracking is used in 23 software packages, 15 of which use GitHub or GitLab
to host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority of their issues closed, and only three (laboetie, lettuce, Sailfish) have less than 50 percent of their issues closed. All of the top five overall ranked packages have most of their issues closed. Alive packages (11 use issue tracking) have 64\% of their issues closed, while dead packages (3 use issue tracking) have 71\% of their issues closed. This information is presented in
Table~\ref{gitrepodata}. With respect to the host for the git repository, 13
packages use GitHub and two (Palabos, waLBerla) use GitLab. Of the other
packages, one package (SunlightLB) uses CVS for issue tracking and version
control, and seven packages do not appear to use any issue tracking system.

@@

Software package artifacts were reviewed for the identification of a specific
development model, like a waterfall or agile development model, and the presence
of documentation recording the development process and standards. They were also
reviewed for the identification of the development environment, and the presence
of release notes. The packages tended to not explicitly use well-known
development models. This was also noted in the interviews with developers
(Section~\ref{Sec_CompareMethodologies}). The development teams of these
packages are fairly small and easily organized without the need for such
processes. Seven of the software packages did have some artifacts outlining the
general development process, how to contribute, and the status of the package or
its components. Eight of the packages have artifacts that note the development
environment. While this information could help developers, and would improve
transparency, the small close-knit nature of the development teams make
explicitly specifying this information practically unnecessary. Version
release notes were found in nine of the software packages.
